---
title: "Stawberries"
author: MA615
date: 2023 Sept 25
format: html
engine: knitr
---

# Preparing data for analysis

## Introduction: foundations

Before we begin to work with the strawberry data, let's talk about how we will approach the work.

### Data cleaning and organization

Cleaning and organizing data for analysis is an essential skill for data scientists.
Serious data analyses must be presented with the data on which the results depend. The credibility of data analysis and modelling depends on the care taken in data preparation and organization.

#### References

In their handbook ["An introduction to data cleaning with R" by Edwin de Jonge and Mark van der Loo](https://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf), de Jonge and van der Loo go into detail about specific data cleaning isssues and how to handle them in R.

["Problems, Methods, and Challenges in Comprehensive Data Cleansing" by Heiko Müller and Johann-Christoph Freytag](https://www.researchgate.net/profile/Heiko-Mueller/publication/228929938_Problems_methods_and_challenges_in_comprehensive_data_cleansing/links/09e415101b58541e2c000000/Problems-methods-and-challenges-in-comprehensive-data-cleansing.pdf) is a good companion to the de Jonge and van der Loo handbook, offering additional issues in their discussion.

### Attitudes

Mechanistic descriptions of data cleaning methods are insufficient

#### Data is the product (or by-product) of purposeful human activity

Much of the data used in analysis accessed on local databases or online which may create the impression that the data have been carefully curated. Beware. Data are produced by people for a purpose, with a point-of-view, and at a time and location that may affect the data. The provenance and lineage of the data are meta data you should include when reporting analysis. Data collection is purposeful human activity with all of the risks and weaknesses that are part of any purposeful human activity.

#### Data is language

Data has meaning. Data can be included in sentences related to the meaning of the data. Cleaning and organizing data should be informed by the meaning the data convey and how that meaning relates to the research you are doing do achieve this important result

-   Immerse yourself in the data. Put data into context.

-   Visualize the data to find problems, confirm your understandings, and plan your data organization. People do a bad job of seeing meaningful patterns in data but a good job of seeing patterns of all kinds when data are rendered as plots. As you product and show visualizations, ask your self and those who view your presentations, "what do you see?" and "what do you wonder?"

## Example: Strawberries

### Public information

[WHO says strawberries may not be so safe for you--2017March16](https://med.news.am/eng/news/13621/who-says-strawberries-may-not-be-so-safe-for-you.html)

[Pesticides + poison gases = cheap, year-round strawberries 2019March20](https://www.ewg.org/foodnews/strawberries.php)

[Multistate Outbreak of Hepatitis A Virus Infections Linked to Fresh Organic Strawberries-2022March5](https://www.cdc.gov/hepatitis/outbreaks/2022/hav-contaminated-food/index.htm)

[Strawberry makes list of cancer-fighting foods-2023May31](https://issuu.com/mechlocal/docs/053123_mech_asf/s/25386339)

## What is the question?

-   Where they are grown? By whom?

-   Are they really loaded with carcinogenic poisons?

-   Are they really good for your health?

-   Are organic strawberries carriers of deadly diseases?

-   When I go to the market should I buy conventional or organic strawberries?

## The data

The data set for this assignment has been selected from: [USDA_NASS_strawb_2023SEP19](https://quickstats.nass.usda.gov) <br> The data have been stored on NASS here: [USDA_NASS_strawb_2023SEP19](https://quickstats.nass.usda.gov/results/45FBC825-B104-38E2-9802-839F5F3C7036)

and has been stored on the backboard as strawberry.csv.

## USDA NASS

```{r}
#| label: Load libraries
#| echo: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
```

## Read the file

```{r warning=FALSE, message=FALSE}
#| label: read data - glimpse 
#| warning: false
#| message: false
#| echo: true

strawberry <- read_csv("strawberry.csv", col_names = TRUE)

glimpse(strawberry)
```


## Remove columns with a single value in all columns

```{r}
#| label: drop one-item columns
#| echo: false

## define function
drop_one_value_col <- function(df){
drop <- NULL
for(i in 1:dim(df)[2]){
if((df |> distinct(df[,i]) |> count()) == 1){
drop = c(drop, i)
} }

if(is.null(drop)){return("none")}else{

   print("Columns dropped:")
   print(colnames(df)[drop])
   strawberry <- df[, -1*drop]
   }
}


## use the function

strawberry <- drop_one_value_col(strawberry)

## applying the function a second time 
## tests the function when there aren't any 
## one-value columns
#####  drop_one_value_col(strawberry)

```



## Examine the data. How is it organized?

### Is every line associated with a state?

```{r}
#| label: examine rows
#| echo: true

## is every line associated with a state?

## state_all1 contains the number of rows containing data 
## for each of the 47 strawberry-growing states.
state_all1 <- strawberry |> group_by(State) |> count()

## test if every row is associated with a state by summing the 
## counts and testing for equality with the total rows in the 
## data frame

if(sum(state_all1$n) == dim(strawberry)[1]){print("Every row has value in the State column.")}

```

### Which state has the most rows?


```{r}
#| label: which state has the most rows
#| echo: true

state_max <- state_all1$State[which(state_all1$n ==  max(state_all1$n)  )]

```



The data is organized by state.  
The state with the most rows is `r paste(state_max)`.

## Examine California data

```{r}
#| label: examine California data
#| echo: true

## filter rows of California data from the CENSUS data
calif_census <- strawberry |> filter((State=="CALIFORNIA") & (Program=="CENSUS"))


## ## filter rows of California data from the SURVEY data
calif_survey <- strawberry |> filter((State=="CALIFORNIA") & (Program=="SURVEY"))

 
census_col <- colnames(calif_census)

survey_col <- colnames(calif_survey)


```



### List the composite columns 

Census: `r paste(census_col[c(6, 8)])`

Survey: `r paste(survey_col[c(6,7,8)])`


## Separate CENSUS and SURVEY into two Data Frames

In the strawberry data frame, 
The CENSUS rows contains marketing, sales, and productiong data.  The SURVEY rows contain rows which may be redundant with the CENSUS rows and chemical application rows.

After splitting CENSUS and SURVEY rows into two data frames, 
finish organizing the CENSUS data first.  Then, organize the
SURVEY data frame splitting the marketing, and production data from the chemical application data.




```{r}
#|label: split srawberry into census and survey pieces

strwb_census <- strawberry |> filter(Program == "CENSUS")

strwb_survey <- strawberry |> filter(Program == "SURVEY")

## check that all of the rows are accounted for

nrow(strawberry) == (nrow(strwb_census) + nrow(strwb_survey))

## Move marketing-related rows in strw_b_chem 
## to strw_b_sales

```


# Complete with the census data frame

## Separate composite columns and clean the Value column

Composite columns in the strwb_census: Data Item, Domain category

Column separators in CENSUS: ",", "-", ":"

### Separate Data Item into columns by ","


```{r}
#| label: split Data Item
#| echo: false

  strwb_census <- strwb_census |>
  separate_wider_delim(  cols = `Data Item`,
                         delim = ",",
                         names = c("Fruit",
                                 "temp1",
                                 "temp2",
                                 "temp3"),
                         too_many = "error",
                         too_few = "align_start"
                       )

## Use too_many and too_few to set up the separation operation.

## split temp1 into crop_type, measure

strwb_census <- strwb_census |>
  separate_wider_delim(  cols = temp1,
                         delim = " - ",
                         names = c("crop_type",
                                 "measure"),
                         too_many = "error",
                         too_few = "align_start"
                       )




## split temp2 into market_type, measure

strwb_census <- strwb_census |>
  separate_wider_delim(  cols = temp2,
                         delim = " - ",
                         names = c("market_type",
                                 "measure2"),
                         too_many = "error",
                         too_few = "align_start"
                       )

## trim the strings

strwb_census$market_type <- str_trim(strwb_census$market_type, side = "both")


## create a vector of Measurement entries from
## the market_type column and put it into the dataframe

# v1 <- strc$market_type[1:30]
# s1 <- "MEA"

combo_1 <- function(v,starts ){
  d2 = str_starts(v, starts)
  d2[is.na(d2)] = FALSE
  out = rep(NA, length(d2))

  for(i in seq_along(d2)){
    if(d2[i]){out[i]=v[i]}
  }
  return(out)
}

start <- "MEA"
new <- combo_1(strwb_census$market_type, start )

strwb_census <- strwb_census |> mutate(measure3 = new, .before = temp3)

```

There is a problem you have to fix -- a leading space.

```{r}
#| label: fix the leading space
#| echo: false

 # note
strawberry$Category[1]
# strawberry$Item[2]
# strawberry$Metric[6]
# strawberry$Domain[1]
## 
## trim white space 

# strawberry$Category <- str_trim(strawberry$Category, side = "both")
# strawberry$Item <- str_trim(strawberry$Item, side = "both")
# strawberry$Metric <- str_trim(strawberry$Metric, side = "both")

```





## The Value column contains 

```{r}
#| label: define functions dcomma and footnote finder
#| echo: true

### remove commas from Value entries
dcomma <- function(c){
  x_new <- as.numeric(gsub(",", "", c))
  return(x_new)
}

## function test
# x1 <- "123,456"
# x2 <- "34"
# x3 <- 45
# x4 <- "  3,456  "
# str(dcomma(x1))
# str(dcomma(x2))
# str(dcomma(x3))
# str(dcomma(x4))
# x5 <- c(x1, x2, x3, x4)
# str(dcomma(x5))
# x6 <- list(x1, x2, x3, x4)
# str(dcomma(x6))


#### finds footnotes
## finds single uppor case Character in parens in s2
## e.g. "(D)"

# d1 <- str_detect(s2, "^\\([:upper:]\\)$") ## returns 
                                        ## vector of locigals
# s2[which(d1)]  ## returns 
               ## the footnotes found


## test
###  footnotes in the Value column
# s2 <- calif_census$Value[1:10] 
# s2[which(d1)]
# s2[6] <- "(2)"   ### fill in test values here
# d1 <- str_detect(s2, "^\\([:upper:]\\)$")
# d1
# s2[which(d1)]

```




### `Data Item` into (fruit, category, item)

```{r}
#|label: split Data Item

  strawberry <- strawberry |>
  separate_wider_delim(  cols = `Data Item`,
                         delim = ",",
                         names = c("Fruit",
                                 "Category",
                                 "Item",
                                 "Metric"),
                         too_many = "error",
                         too_few = "align_start"
                       )

## Use too_many and too_few to set up the separation operation.

```

There is a problem you have to fix -- a leading space.

```{r}
#|label: fix the leading space

 # note
strawberry$Category[1]
# strawberry$Item[2]
# strawberry$Metric[6]
# strawberry$Domain[1]
## 
## trim white space 

strawberry$Category <- str_trim(strawberry$Category, side = "both")
strawberry$Item <- str_trim(strawberry$Item, side = "both")
strawberry$Metric <- str_trim(strawberry$Metric, side = "both")

```

## now exam the Fruit column -- find hidden sub-columns

```{r}

unique(strawberry$Fruit)

## generate a list of rows with the production and price information

spr <- which((strawberry$Fruit=="STRAWBERRIES - PRODUCTION") | (strawberry$Fruit=="STRAWBERRIES - PRICE RECEIVED"))


strw_prod_price <- strawberry |> slice(spr)


## this has the census data, too

strw_chem <- strawberry |> slice(-1*spr)  ## too soon

```

## now examine the rest of the columns

Which ones need to be split?



## plots

```{r}
#|label: plot 1

plot1_data <- strawberry |> 
  select(c(Year, State, Category, Value)) |> 
  filter((Year == 2021) & (Category == "ORGANIC - OPERATIONS WITH SALES"))

plot1_data$Value <- as.numeric(plot1_data$Value)

plot1_data <- plot1_data |> arrange(desc(Value))

ggplot(plot1_data, aes(x=reorder(State, -Value), y=Value)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x = "States", y = "Count",
title ="Number of Organic Strawberry operations with Sales in 2021")


```

```{r}
## plot 2

plot2_data <- strawberry |> 
  select(c(Year, State, Category, Item, Value)) |> 
  filter((Year == 2021) & 
           (Category == "ORGANIC - SALES") & 
           (Item == "MEASURED IN $") & 
           (Value != "(D)"))


plot2_data$Value <- as.numeric(gsub(",", "", plot2_data$Value))

plot2_data <- plot1_data |> arrange(desc(Value))

ggplot(plot2_data, aes(x=reorder(State, -Value), y=Value)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x = "States", y = "Sales",
title ="Organic Strawberry Sales ($) in 2021")


```

## chemicals

### carcinogens from WHO list

Six specific pesticides - captafol, ethylene dibromide, glyphosate, malathion, diazinon and dichlorophenyltrichloroethane (DDT)

```{r}

cc <- strawberry |> distinct(Category)

cca <- strawberry |> 
       distinct(Domain)

## Split domain into two columns "type" and "subtype"
  strawberry <- strawberry |>
  separate_wider_delim(  cols = Domain,
                         delim = ",",
                         names = c("type",
                                 "subtype"),
                         too_many = "error",
                         too_few = "align_start"
                       )
         
## check the result
    
  ctype <- strawberry |> distinct(type)

  
  csubtype <- strawberry |> distinct(subtype)
  
## 
  
  
  
##  
  
  yr <- strawberry |> distinct(Year)
  

```

```{r}
## columns need descriptive names

doc_cat <- strawberry |> distinct(`Domain Category`)

strawberry <- strawberry |>
  separate_wider_delim(  cols = `Domain Category`,
                         delim = ",",
                         names = c("type1",
                                 "detail1",
                                 "detail2",
                                 "datail3"),
                         too_many = "error",
                         too_few = "align_start"
                       )

```

```{r}
## columns need descriptive names

strawberry <- strawberry |>
  separate_wider_delim(  cols = type1,
                         delim = ":",
                         names = c("type1a",
                                 "type1b"),
                         too_many = "error",
                         too_few = "align_start"
                       )


```

for EPA number lookup [epa numbers](https://archive.epa.gov/pesticides/chemicalsearch/chemical/foia/web/html/128810.html)

[Active Pesticide Product Registration Informational Listing](https://ordspub.epa.gov/ords/pesticides/f?p=APPRIL_PUBLIC:2::::::)

[CAS for Methyl Bromide](https://ordspub.epa.gov/ords/pesticides/f?p=113:17::::::)

[pesticide chemical search](https://ordspub.epa.gov/ords/pesticides/f?p=chemicalsearch:1)

[toxic chemical dashboard](https://comptox.epa.gov/dashboard/)

[pubChem](https://pubchem.ncbi.nlm.nih.gov/)

The EPA PC (Pesticide Chemical) Code is a unique chemical code number assigned by the EPA to a particular pesticide active ingredient, inert ingredient or mixture of active ingredients.

```{r}

# dat1 <- strawberry |> filter(type=="CHEMICAL")
# 
# dat2 <- strawberry |> filter(strawberry$type!=strawberry$type1a)
# 
#  
# data_f21 <- strawberry |> 
#   filter((subtype == " FUNGICIDE") & (State == "CALIFORNIA") & (Year == "2021")  )
# 
# data_f20 <- strawberry |> 
#   filter((subtype == " FUNGICIDE") & (State == "CALIFORNIA") & (Year == "2020")  )
# 
# data_f19 <- strawberry |> 
#   filter((subtype == " FUNGICIDE") & (State == "CALIFORNIA") & (Year == "2019")  )
# 
# data_f18 <- strawberry |> 
#   filter((subtype == " FUNGICIDE") & (State == "CALIFORNIA") & (Year == "2018")  )
# 
# data_f17 <- strawberry |> 
#   filter((subtype == " FUNGICIDE") & (State == "CALIFORNIA") & (Year == "2017")  )
# 



```

```{r}
## columns need descriptive names

strawberry <- strawberry |>
  separate_wider_delim(  cols = detail1,
                         delim = ":",
                         names = c("detail1a",
                                 "detail1b"),
                         too_many = "error",
                         too_few = "align_start"
                       )



strawberry$detail1b <- strawberry$detail1b |> 
  str_trim(side = "both") |> 
     str_sub(start = 2, end = -2)


aa <- strawberry$detail1b

aa <- na.omit(aa)


group1 <- c("captafol", "ethylene dibromide", 
            "glyphosate","malathion", "diazinon", 
            "dichlorophenyltrichloroethane", "DDT")


```

Investigating toxic pesticides

[start here with chem PC code](https://ordspub.epa.gov/ords/pesticides/f?p=chemicalsearch:1)

[step 2](https://ordspub.epa.gov/ords/pesticides/f?p=113:1::::RP,17,1::) to get label (with warnings) for products using the chemical

[Pesticide Product and Label System](https://ordspub.epa.gov/ords/pesticides/f?p=113:1::::RP,17,1::)

[Search by Chemical](https://ordspub.epa.gov/ords/pesticides/f?p=113:17::::::)

[CompTox Chemicals Dashboard](https://comptox.epa.gov/dashboard/)

[Active Pesticide Product Registration Informational Listing](https://ordspub.epa.gov/ords/pesticides/f?p=APPRIL_PUBLIC:2::::::)

[OSHA chemical database](https://www.osha.gov/chemicaldata)

[Pesticide Ingredients](http://npic.orst.edu/ingred/)

[NPIC Product Research Online (NPRO)](http://npic.orst.edu/NPRO/)

[Databases for Chemical Information](http://npic.orst.edu/ingred/cheminfo.html)

[Pesticide Active Ingredients](http://npic.orst.edu/ingred/active.html)

[TSCA Chemical Substance Inventory](https://www.epa.gov/tsca-inventory)

[glyphosate](https://ordspub.epa.gov/ords/pesticides/f?p=CHEMICALSEARCH:3::::1,3,31,7,12,25:P3_XCHEMICAL_ID:2478)
